---
title: "10/14 Notes"
author: "Sepehr Akbari"
date: "10/14/2024"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
```


## Unite

In table5, the year variable is split between two columns.

```{r}
table5

table5 %>% 
  unite(
    col = "year",
    century, year,
    sep = "") # don't seperate the parts of the new string
```

## Joining data sets

It's common to want to add column from one data project to another.

```{r}
band_members

band_instruments
```

We do this with `left_join`. We need to specify both datasets starting with the one we want to modify. We also need to specify one or more **key columns**, ones that identify observations, letting R know how to match up the rows.

```{r}
band_members %>%
  left_join(band_instruments, 
            by = "name")
```

By default, `left_join` adds all columns in the second set to th firstm with the exception of the key columns.

You can also use `rbind` and `cbind` to just combine sets by row or column. For instance if you have two sets with identical variables and different observations, use `rbind.`
