---
title: "10/01 Notes"
author: "Sepehr Akbari"
date: "10/01/2024"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r}
#| warning: FALSE
#| message: FALSE

library(tidyverse)
library(modeldata)
```

# A few things with `penguins`

First, let's do grouped regression lines.

```{r}
#| warning: FALSE
#| fig-alt: "(description here)" # alt text

ggplot(
  penguins,
  aes(
    x = bill_length_mm,
    y = bill_depth_mm,
    color = species
  )
) +
  geom_point() +
  geom_smooth( # regression line(s)
    se = FALSE, # hiding the error ribbon
    method = "lm") + # making the lines linear
  scale_color_brewer(palette = "Dark2") +
  theme_minimal()
```

Next let's get the 3 penguins of each specie with the shortest bills:

```{r}
penguins %>%
  group_by(species) %>%
  slice_min(order_by = bill_length_mm, n = 3)
```

Notice that `slice_min` respected grouping. The same is true of `filter` and `mutate`. For instance:

```{r}
peguins_new <- penguins %>%
  group_by(species) %>%
  mutate(avg_bills = mean(bill_length_mm, na.rm = TRUE)) %>% 
  select(species, bill_length_mm, bill_depth_mm, avg_bills)
```

Let's look at this in the viewer. We only have 3 unique values in the `avg_bill` column.

## Inline code

The average bill length in this set is `{r} mean(penguins$bill_length_mm, na.rm = TRUE)`.








