---
title: "10/11 Notes"
author: "Sepehr Akbari"
date: "10/11/2024"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

## HW 10/11

```{r}
#| label: hidden
#| message: FALSE
#| warning: FALSE
#| error: FALSE
#| echo: FALSE
base_dir <- "/Users/sepehrakbari/Documents/LFC/Semester 3/MATH 250/MATH_250"
```

```{r}
#| label: setup
#| message: FALSE
#| warning: FALSE
#| error: FALSE

library(tidyverse)
library(readxl)
library(gt)

airport_data_speeds <- read_excel(file.path(base_dir, "Data/airport_data_speeds.xlsx"))
```

#### Problem 2

```{r}
speeds_long <- airport_data_speeds %>%
  pivot_longer(
    cols = VERIZON:`T-MOBILE`,
    names_to = "carrier",
    values_to = "speed"
  )
speeds_long %>% 
  slice_head(n = 5) %>% 
  gt::gt()
```

#### Problem 3

We want a table showing how many airports have each carrier as the fastest.

```{r}
speeds_long %>% 
  group_by(`AIRPORT CODE`) %>% 
  filter(speed == max(speed)) %>% 
  ungroup() %>%
  count(carrier) %>% 
  arrange(desc(n)) %>% 
  gt::gt()
```

## Pivot Wider

1.  Variables are spread across multiple columns.

```{r}
glimpse(table2)
```

Here each observation has multiple rows. We fix this with `pivot_wider`, which is basically the reverse of `pivot_longer`.

```{r}
table2_wide <- table2 %>% 
  pivot_wider(
    id_cols = c(country, year),
    names_from = type,
    values_from = count
  )

gt::gt(table2_wide)
```

2.  Variables are combined in one column.

```{r}
glimpse(table5) # rate
```

Use the separate command (or one of its friends) to split the `rate` column into population and cases.

```{r}
table5 %>% 
  separate(
    col = rate,
    into = c("cases", "population")
  ) %>% 
  gt::gt()
```

Use `sep = "/"` to specify the separator. Otherwise R will guess.

3.  A single variable is split between multiple columns

```{r}
glimpse(table5) # century and year
```

Use `unite` to handle this.
