---
title: "10/4 Notes"
author: "Sepehr Akbari"
date: "10/04/2024"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r}
#| message: FALSE
s
library(tidyverse)
library(readxl)
library(gt)

astronauts <- read_excel(file.path(base_dir, "Data/astronauts.xlsx"))
```

## Astronauts HW Problem 2 (top occupations)

Rough and ready:

```{r}
table(astronauts$occupation)
```

Notice the inconsistent coding. We should spend some time on this, ideally. Let's start by ignoring it though, since the numbers are small.

Let's be a bit more methodical about our counting.

```{r}
astronauts %>% 
  group_by(occupation) %>%
  summarise(count = n()) %>% 
  slice_max(order_by = count, n = 5) %>%
  gt()
```

You could also use `count()` instead of `group_by()` and `filter()` instead of `slice_max()`.

Let's improve the `occupation` column a little with a mutate.

```{r}
astronauts <- astronauts %>% 
  mutate(occupation = tolower(occupation))

table(astronauts$occupation)
```

A useful tool for more fine grained recoding of categorical data is `case_when`. 

## Problem 3

```{r}
mil_civ_year <- astronauts %>% 
  group_by(year_of_mission, military_civilian) %>%
  summarise(eva_tot = sum(eva_hrs_mission)) %>%
  ungroup()

mil_civ_year %>%
  slice_head(n = 10) %>% 
  gt()
```

## Tidy Data

compare the following

```{r}
table1
table4a
table4b
```

In the last two, some of the column names are actually values of an unnamed variable. In 4a and 4b, the variables have been divided between the two sets.

```{r}
table2
```








