---
title: "10/7 Notes"
author: "Sepehr Akbari"
date: "10/04/2024"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(gt)
```

```{r}
gt(table4a)
```


In a tidy data set:

- every column is one variable

- every row is one observation

- there is only one kind of observational

The `table4a` dataset does not obey the first two principles. Here, the year variable appears as column names and the cases variable appears as cell values in multiple columns. This is a common untidy structure **wide data**. We tidy such set with `pivot_longer`, which requires 3 pieces of information.

1) the problem columns

2) the name(s) of the new column(s) where the problem column names will be housed. In this case, "year."

3) the names of the new column(s) where the corresponding values will go. In this case, "cases."
```{r}
table_long <- table4a %>% 
  pivot_longer(
    cols = `1999`:`2000`,
    names_to = "year",
    values_to = "cases"
  )

gt(table_long)
```

*Let's do a fuller example:*

## The billboard hot 100

This data does not print nicely, but let's see it anyway.

```{r}
gt(billboard)
```

We'd like a variable called "week" and another called "rank." This is a pivoting problem. Since the names of columns are actually values of the week variable, this is a `pivot_longer` situation.

```{r}
billboard_long <- billboard %>% 
  pivot_longer(
    cols = contains("wk"),
    names_to = "week",
    values_to = "rank"
  )

gt(billboard_long)
```

The help file is, well, helpful. So check it out with `?pivot_longer`.

```{r}
billboard_long <- billboard %>% 
  pivot_longer(
    cols = contains("wk"),
    names_to = "week",
    values_to = "rank",
    names_prefix = "wk",
    values_drop_na = TRUE
  )

gt(billboard_long)
```

Now we need to make `week` numerical.

```{r}
billboard_long <- billboard_long %>% 
  mutate(
    week = as.numeric(week)
  )

gt(billboard_long)
```

Also check out `as.charachter`, and others.

Next time: 

- What are the top 3 songs?

- How can we go from "week" to "date"?

- Make a great plot!






