---
title: "9/11 Notes"
author: "Sepehr Akbari"
date: "9/11/2024"
format: html
embed-resources: true
---

```{r}
#| message: false
#| label: setup

library(tidyverse)
library(readxl)
scooby <- read_excel(file.path(base_dir, "Data/scooby.xlsx"))
```


## Scooby and `imdb`

Last time we looked at the `imdb` variable alone. Now let's consider a second.

Let's start with time:

```{r}
#| label: time & imdb

ggplot(
  scooby,
  aes(
    x = date_aired,
    y = imdb
  )
) +
  geom_point()
```

What's goind on around 2008-2009? Does it have to do with series name or network?

```{r}
ggplot(
  scooby,
  aes(
    x = date_aired,
    y = imdb,
    color = series_name
  )
) +
  geom_point()
```

This plot is cluttered and unclear.

First let's deal with those missing values and eliminate the warning message. We have seen that the most recent episodes don't have `imdb` rating yet. We comment on this in our report and then suppress R's warning message.

*Principle*: having lots of categories in a plot is generaly not clear. Limit use of color to six if possible and four if you can.

Let's stop considering time, and instead just look at imdb vs series and vs network.

```{r}
#| label: imdb & series
#| warning: false

ggplot(
  scooby,
  aes(
    x = imdb,
    y = series_name
  )
) +
  geom_boxplot()
```

Let's try network:

```{r}
#| label: imdb & network
#| warning: false

ggplot(
  scooby,
  aes(
    x = imdb,
    y = network
  )
) +
  geom_boxplot()
```

It looks like *The CW's "Shaggy & Scooby-Doo Get a Clue!"* was hated. We should read about it.

