---
title: "9/13 Notes"
author: "Sepehr Akbari"
date: "9/13/2024"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(readxl)

scooby <- read_excel(file.path(base_dir, "Data/scooby.xlsx"))
```

Let's start subsetting our data.

## Choosing Rows

We subset rows using criteria that we specify using the `filter` command from the `dplyer` package, which loads with `tidyverse`. 

For instance, let's say we just want episodes on CBS.

```{r}
#| label: filtering CBS network

scooby_cbs <- filter(
  scooby,
  network == "CBS"
)
```

The double equal sign means 'logical equality'. Compare with "=" which specifies the name of an argument.

R does NOT automatically save this, or any other result.

Notice that there are fewer rows in `scooby_cbs` than in `scooby`. But the same number of columns.

```{r}
nrow(scooby_cbs)
ncol(scooby_cbs)
```

We can filter numarically.

```{r}
#| label: filtering by imdb rating

scooby_good <- filter(
  scooby,
  imdb >= 7.0
)

scooby_better <- filter(
  scooby,
  imdb >= mean(imdb,
               na.rm = TRUE)
)
```

We can filter for multiple conditions at once.

```{r}
#| label: filtering by network & season

scooby_cbs_s1 <- filter(
  scooby,
  network == "CBS",
  season == 1
)
```

If we want everything but CBS:

```{r}
#| label: filtering out CBS

scooby_not_cbs <- filter(
  scooby,
  network != "CBS"
)
```

What about missing values? Most machine learning models treat missing data as random, which is normally not.

```{r}
#| label: filtering by missing values (incorrect)

scooby_NA <- filter(
  scooby,
  imdb == "NA"
)
```

Here, R thinks I'm looking for `imdb` values that are equal to the string "NA". Instead, I want to look for missing values. To detect missing values, we use the `is.na` function.

```{r}
#| label: filtering by missing values

scooby_NA <- filter(
  scooby,
  is.na(imdb)
)
```

## Old-fashioned Subsetting

We can subset vectors using brackets.

```{r}
#| label: subsetting vectors

my_vec <- c(1, 7, -3, 4, 10, 15)

# ex, for third element 
my_vec[3]

# ex, for third to fifth elements
my_vec[3:5]

# ex, removing the third element
my_vec[-3]

# ex, removing the third to fifth elements
my_vec[-(3:5)]
```

For a data frame, we have to specify row and column.

```{r}
scooby[1, 2:4] # row1, columns 2-4
```

We almost never need to use this in data anlysis. `filter` and  `select` are much more useful.







