---
title: "9/16 Notes"
author: "Sepehr Akbari"
date: "9/16/2024"
format: html
embed-resources: true
---

## Considering the `scat` homework

Some things I'll be looking for:

- Is the code valid and appropriate to this class and to 2024?

- Are the variables used explained and/or contextualized?

- Is there writing that supports the code ad which is helpful to the reader?

- Does each plot communicate a specific data idea or answer?

- Does the quarto document look good?

## Using `filter` and `select`.

Let's work with `scat` a bit more.

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(modeldata)
```

Suppose we want a subset with bobcats or coyotes.

```{r}
scat_bc <- filter(
  scat,
  Species == "bobcat" | Species == "coyote"
)

glimpse(scat_bc)
```

For more, check out the `filter` video.

Let's look at subsetting columns. This means selecting variables of interest and omitting the others.

```{r}
scat_sm <- select(
  scat,
  Species,
  Number,
  Age,
  Length
)

glimpse(scat_sm)
```

Two other ways to the same thing:

```{r}
scat_sm <- select(
  scat,
  1, 6, 7, 8 
) # Not Recommended

scat_sm <- select(
  scat,
  Species,
  Age:Length
)
```

You might want to keep columns that start with letter "S".

```{r}
scat_sm <- select(
  scat,
  starts_with("S")
)

glimpse(scat_sm)
```

We can remove a column:

```{r}
scat_sm <- select(
  scat,
  -(ropey:scrape)
)

glimpse(scat_sm)
```

Other times, you might just want to move columns around.

```{r}
scat_new <- select(
  scat,
  Location,
  Species,
  everything()
)

glimpse(scat_new)
```

The `everything` helper says to include all the remaining columns after the specified ones.

## Next time: re-arranging rows

We'll use the `arrange` command.

