---
title: "9/27 Notes"
author: "Sepehr Akbari"
date: "9/27/2024"
format: html
embed-resources: true
---

## Making great looking tables in Quarto

We'll do this using the `gt` package, which we can install with: `install.packages("gt")`.

```{r}
#| message: FALSE
#| warning: FALSE
library(tidyverse)
library(modeldata)
library(readxl)
library(gt)

scooby <- read_excel(file.path(base_dir, "Data/scooby.xlsx"))
```

In the package `gt`, the function `gt` just makes a nice table out of data frames. Let's use this to make a nice table of mean bill lengths for different species of penguins.

```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(avg_bill = mean(bill_length_mm, na.rm = TRUE))

# ok... but not great looking
```


```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(avg_bill = mean(bill_length_mm, na.rm = TRUE)) %>% 
  gt()

# much better!
```

## Some group summery practice with scooby

Did scooby use his catchphrase more over the years? Less? Or was he consistent?

Let's start with a rough explaratory plot

```{r}
ggplot(
  scooby,
  aes(
    x = date_aired,
    y = rooby_rooby_roo
  )
) +
  geom_point()
```

We should resolve the warning. Let's not dwell on this too much, so let's do it fast. 

```{r}
scooby_missing <- scooby %>%
  filter(is.na(rooby_rooby_roo))
  #glimpse()
table(scooby_missing$series_name)
```

Let's just drop these. They're all the same show, which we should consider separately.

```{r}
scooby <- scooby %>%
  filter(!is.na(rooby_rooby_roo))
```


By the way, `drop_na()` is a good option for this.

New idea: group by year and get mean scoobies per episode every year.
