---
title: "9/30 Notes"
author: "Sepehr Akbari"
date: "9/30/2024"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: FALSE

library(tidyverse)
library(readxl)

scooby <- read_excel(file.path(base_dir, "Data/scooby.xlsx"))
```


Let's try to get a better plot for Scooby's Roobies over time. Let's try to get the average number of roobies per year in each year.

```{r}
rooby_by_year <- scooby %>% 
  group_by(year) %>% 
  summarize(avg_roobies = mean(rooby_rooby_roo, na.rm = TRUE),
            num_episodes = n())
```

Let's just print a decent looking table of this, using the `gt()` function from the `gt` package.

```{r}
gt::gt(rooby_by_year) # package::function without using library(package)
```


How can we make this shorter, either just taking the first few rows or maybe just taking the years with the most roobies?

```{r}
ggplot(
  rooby_by_year,
  aes(
    x = year,
    y = avg_roobies,
    size = num_episodes
  )) +
  geom_point() +
  geom_smooth()
```

Let's fix the warning, remove the error ribbon, and add a minimal theme.

```{r}
ggplot(
  rooby_by_year,
  aes(
    x = year,
    y = avg_roobies
  )) +
  geom_point(aes(size = num_episodes)) +
  geom_smooth(se = FALSE) +
  theme_minimal()
```

## Alt text

We add descriptive text to plots by using the `fig-alt` option. Please get in this habit. It is graded!

```{r}
#| fig-alt: A scatterplot showing roobies per year with size representing number of episodes. The trend is an upside down U.

ggplot(
  rooby_by_year,
  aes(
    x = year,
    y = avg_roobies
  )) +
  geom_point(aes(size = num_episodes)) +
  geom_smooth(se = FALSE) +
  theme_minimal()
```

## Slicing a data set

A few smaller useful `tidyverse` commands.

```{r}
slice_head(rooby_by_year, n = 5)
```

Let's get the 5 year with the most avg roobies.

```{r}
rooby_by_year %>% 
  slice_max(order_by = avg_roobies, n = 5)
```


























