---
title: "9/4 Notes"
author: "Sepehr Akbari"
date: "9/4/2024"
format: html
embed-resources: true
---

## Intro

Here is some text. Here is *italic* and **bold**, and a footnote ^[like this]. If you want something to typset looking like code, use ticks. `library(tidyverse)`.

```{r}
library(tidyverse) 
```

How can we set output options to supress the message or change how plot looks?

By the way, Quarto is language neutral:

```{python}
3 + 7
```

## Bar Charts

We'll be using the `Carsets` from the `ISLR` package. We haven't installed this yet.

```{r}
library(ISLR2)
glimpse(Carseats)
```

You can learn more about the set with `?Carseats`, but don't do that in Quarto (do it in console).

**Question**: How are `ShelveLoc` and `US` relates, if at all?

```{r}
ggplot(
  Carseats,
  aes(
    x = ShelveLoc,
    fill = US
  )
) +
  geom_bar() +
  scale_fill_brewer(palette = "Dark2")
```

A stacked bar chart emphasizes the variable on the x-axis. Let's compare with this:

```{r}
ggplot(
  Carseats,
  aes(
    fill = ShelveLoc,
    x = US
  )
) +
  geom_bar() +
  scale_fill_brewer(palette = "Dark2")
```

Sometimes we are more interested in the relative perportions of the secondary (fill) variable. So let's do this:

```{r}
ggplot(
  Carseats,
  aes(
    fill = ShelveLoc,
    x = US
  )
) +
  geom_bar(position = "fill") +
  scale_fill_brewer(palette = "Dark2")
```

But now we can't see the amount... So let's try this maybe:

```{r}
ggplot(
  Carseats,
  aes(
    fill = ShelveLoc,
    x = US
  )
) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Dark2")
```

This is a good chart, but its actually the least useful.