---
title: "9/9 Notes"
author: "Sepehr Akbari"
date: "9/9/2024"
format: html
embed-resources: true
---

## Troubkeshooting

What is your progress when you get an error?

- Read the error message carefully. R messages are often very specific. Quarto rendering errors for instance, include the line numbers where the problem is.

- Check spelling and syntax.

- AI (like GPTs) can be really useful for finding syntax and/or logical errors sometimes.

- Read about the problem function or tool.

- Restarting R.

- Ask someone.

## Projects & file management

By now, you probably have a folder for this class. We like to identify that folder as representing the home for a particular data science project (set of related tasks). We'd like R to now it's where all our related files live.

We do this with an **RStudio Project**.

When you work inside a project, data import is much easier. 

```{r}
library(readxl)
scooby <- read_excel(file.path(base_dir, "Data/scooby.xlsx"))
```

## Scooby and IMDB

Let's explore the ratings of scooby Doo episodes a little.

```{r}
library(tidyverse)
glimpse(scooby)
```

First, let's get a one-variable plot for IMDB, our primary variable of interest. This 

```{r}
ggplot(
  scooby,
  aes(x = imdb)
) +
  geom_histogram()
```

Similarly we should consider numerical summeries.

```{r}
mean(scooby$imdb, na.rm = TRUE)
```

Next, we'll explore relations of other variables with this one. We'll start with  `date_aired` next time.



